---
title: "Financial Time Series Models"
format: html
---

```{r ,echo=FALSE, message=FALSE, warning=FALSE}
library(flipbookr)
library(tidyverse)
library(ggplot2)
library(forecast)
library(astsa) 
library(xts)
library(tseries)
library(fpp2)
library(fma)
library(lubridate)
library(tidyverse)
library(TSstudio)
library(quantmod)
library(tidyquant)
library(plotly)
library(ggplot2)
library(FinTS)
library(fGarch)
```

```{r echo=FALSE}

options("getSymbols.warning4.0"=FALSE)
options("getSymbols.yahoo.warning"=FALSE)

tickers = c("JPM")
for (i in tickers){
  getSymbols(i,
             from = "2012-10-01",
             to = "2022-12-01")}

x <- list(
  title = "date"
)
y <- list(
  title = "value"
)

stock <- data.frame(JPM)
stock <- data.frame(stock,rownames(stock))
stock$date <- as.Date(stock$rownames.stock.,"%Y-%m-%d")
tail(stock)
```

```{r}
stock$date=as.factor(stock$date)
fin=ts(stock$JPM.Adjusted,
            start = decimal_date(as.Date("2012-10-01")),
            frequency = 365.25)
str(fin)
autoplot(fin, main = "Time Series Plot for JPM Stock(2012-2022)")
```

```{r}
log_jpm = log(fin)
returns = diff(log_jpm)
par(mfrow = c(3,1))
plot(returns , col = 'darkgreen', 
     main = 'log return of JPM Closing Price',
     ylab='Closing Price', xlab='Date')
acf(returns, col ="darkgreen")
pacf(returns,col = "darkgreen")


```

```{r}
adf.test(returns)
```

```{r}
arima <- Arima(returns,order=c(1,0,0))
summary(arima) 
```

```{r}
auto.arima(returns) #ARIMA(1,0,0)
```

```{r}
sarima(returns,1,0,0)
```

```{r}
res = arima$residuals

#ARCH test on residual
ArchTestRes <- as.character(ArchTest(res)$p.value)
names(ArchTestRes)<- "JPM"
ArchTestRes

```

Because the p-value is less than 0.05, meaning that we need some other model to fit it since there is ARCH effect in the residuals for the ARMA model.

```{r}
plot(res^2,main = "Squared Residual") 
par(mfrow=c(2,1))
acf(res^2) # q = 1
pacf(res^2) # p = 1，2
```

```{r}
garch.fit1 <- garchFit(formula = ~arma(1,0) + garch(2,1), data = returns, trace=F)
summary(garch.fit1)
```

```{r}
garch.fit2 <- garchFit(formula = ~arma(1,0) + garch(1,1), data = returns, trace=F)
summary(garch.fit2)

```

Because the AIC and BIC of fit2 is lower than that of fit1, so we say the second model performs better, which is arma(1,0) + garch(1,1).

## Summary

After analysis, we fitted an ARMA(1,0)+GARCH(1,1) model for the returns of JPM stock. According to Ljung-Box test, all p-value are greater than 0.05, meaning that there is no correlation left in the residuals so we can say that the model fitted well.

ϕ(B)x_t=δ+θ(B)y_t

ϕ(B)=1−0.003B

var(y_t\|y_t−1) = σ\^2 = 4.19×10\^-5 + 0.16(y\_{t−1})\^2 + 0.74σ\^2\_{t_1}
