---
title: "Financial Time Series Models (ARCH/GARCH)"
format: html
---

## Original Data Visualization

```{r ,echo=FALSE, message=FALSE, warning=FALSE}
library(flipbookr)
library(tidyverse)
library(ggplot2)
library(forecast)
library(astsa) 
library(xts)
library(tseries)
library(fpp2)
library(fma)
library(lubridate)
library(tidyverse)
library(TSstudio)
library(quantmod)
library(tidyquant)
library(plotly)
library(ggplot2)
library(FinTS)
library(fGarch)
```

```{r echo=FALSE}

options("getSymbols.warning4.0"=FALSE)
options("getSymbols.yahoo.warning"=FALSE)

tickers = c("JPM")
for (i in tickers){
  getSymbols(i,
             from = "2012-10-01",
             to = "2022-12-01")}

x <- list(
  title = "date"
)
y <- list(
  title = "value"
)

stock <- data.frame(JPM)
stock <- data.frame(stock,rownames(stock))
stock$date <- as.Date(stock$rownames.stock.,"%Y-%m-%d")
tail(stock)
```

```{r echo=FALSE, message=FALSE}
stock$date=as.factor(stock$date)
fin=ts(stock$JPM.Adjusted,
            start = decimal_date(as.Date("2012-10-01")),
            frequency = 365.25)
str(fin)
autoplot(fin, main = "Time Series Plot for JPM Stock(2012-2022)")
```

Looking at JP Morgan's stock price from 2012 to 2020, we can see that it experienced significant fluctuations in both its stationarity and volatility.

During this period, JP Morgan's stock price generally trended upwards, but there were several instances of major market disruptions that caused significant declines in the stock's value. For example, in early 2016, concerns about a potential recession caused the stock price to drop by around 20% in just a few months. Similarly, in early 2020, the COVID-19 pandemic caused widespread market turmoil, resulting in a nearly 40% drop in the stock's value over a period of just a few weeks.

In terms of stationarity, the mean and variance of JP Morgan's daily returns showed some evidence of non-stationarity during this period. There were several instances where the mean return varied significantly, indicating a lack of stationarity in the stock's behavior. Additionally, the variance of the daily returns was generally higher during periods of market volatility, such as during the 2016 and 2020 market disruptions.

Overall, JP Morgan's stock price from 2012 to 2020 exhibited both stationarity and volatility, with significant fluctuations in both properties over time.

## Model Fitting

```{r echo=FALSE, message=FALSE}
log_jpm = log(fin)
returns = diff(log_jpm)
par(mfrow = c(3,1))
plot(returns , col = 'darkgreen', 
     main = 'log return of JPM Closing Price',
     ylab='Closing Price', xlab='Date')
acf(returns, col ="darkgreen")
pacf(returns,col = "darkgreen")


```

```{r echo=FALSE, message=FALSE}
adf.test(returns)
```

```{r echo=FALSE, message=FALSE}
arima <- Arima(returns,order=c(1,0,0))
summary(arima) 
```

```{r echo=FALSE, message=FALSE}
auto.arima(returns) #ARIMA(1,0,0)
```

```{r echo=FALSE, message=FALSE}
sarima(returns,1,0,0)
```

```{r echo=FALSE, message=FALSE}
res = arima$residuals

#ARCH test on residual
ArchTestRes <- as.character(ArchTest(res)$p.value)
names(ArchTestRes)<- "JPM"
ArchTestRes

```

Because the p-value is less than 0.05, meaning that we need some other model to fit it since there is ARCH effect in the residuals for the ARMA model.

## Additional Model Fitting

```{r echo=FALSE, message=FALSE}
plot(res^2,main = "Squared Residual") 
par(mfrow=c(2,1))
acf(res^2) # q = 1
pacf(res^2) # p = 1，2
```

```{r echo=FALSE, message=FALSE}
garch.fit1 <- garchFit(formula = ~arma(1,0) + garch(2,1), data = returns, trace=F)
summary(garch.fit1)
```

```{r echo=FALSE, message=FALSE}
garch.fit2 <- garchFit(formula = ~arma(1,0) + garch(1,1), data = returns, trace=F)
summary(garch.fit2)

```

Because the AIC and BIC of fit2 is lower than that of fit1, so we say the second model performs better, which is arma(1,0) + garch(1,1).

## Summary

After analysis, we fitted an ARMA(1,0)+GARCH(1,1) model for the returns of JPM stock. According to Ljung-Box test, all p-value are greater than 0.05, meaning that there is no correlation left in the residuals so we can say that the model fitted well.

ϕ(B)x_t=δ+θ(B)y_t

ϕ(B)=1−0.003B

var(y_t\|y_t−1) = σ\^2 = 4.19×10\^-5 + 0.16(y\_{t−1})\^2 + 0.74σ\^2\_{t_1}

## Source Code

[Github](https://github.com/YujieChen20/Time-Series/blob/main/financial.qmd)
